---
title: "Gapminder Analysis in R"
author: "Jinyu Gan"
date: 2026-09-24
---

### Data

This data used in this post come from the [Gapminder dataset](https://www.gapminder.org/data/).

Gapminder dataset contains country-level information including life expectancy, population, and GDP per capita across multiple years.

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(gapminder)
library(dplyr)
library(ggplot2)
```

### Question

Which continent had the highest average life expectancy in 2002?

### First step

We filter the dataset to keep only observations from 2002.

```{r}
#| label: Filter the dataset

year_2002 <- gapminder |> 
  filter(year == 2002)

year_2002
```

### Second Step

We group the 2002 observations by continent and calculate the average life expectancy for each continent.

```{r}
#| label: Find out the average life expectancy
life_exp_2002 <- year_2002 |> 
  group_by(continent) |> 
  summarize(
    mean_life_exp = mean(lifeExp)
  )

life_exp_2002

```

### Last Step

Finally, we identify the continent with the highest average life expectancy in 2002.

```{r}
#| label: identify the continent
continent_ranking <- life_exp_2002 |> 
  arrange(desc(mean_life_exp))

continent_ranking[1,1]
```

### Findings

Based on the average life expectancy values in 2002, Oceania had the highest average life expectancy among the continents in the Gapminder dataset.

```{r}
#| label: fig-life-exp-2002
#| fig-cap: "Average life expectancy by continent in 2002."

ggplot(life_exp_2002, aes(x = continent, y = mean_life_exp)) +
  geom_col()
```

The bar chart also shows that there were clear differences in average life expectancy across continents.
